---
title: "liveassignment11"
author: "frank sclafani"
date: "11/27/2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1)	Go through the electric equipment example. 
# a)	Plot the time series. Can you identify seasonal fluctuations and/or a trend? 

```{r visitors}
library(fpp)
data(visitors)
```

## Plot the data
# Yes, seasional trends are able to be identified by the peaks and valleys occuring on a regular basis.
```{r plot1}
plot(visitors)
```

#b)	Use a classical decomposition to calculate the trend-cycle and seasonal indices. 
```{r decompose}
data(visitors)
decompose(visitors)
```

##c)	Do the results support the graphical interpretation from part (a)? 
Yes, the reoccuring values in the seasonal data ser support the gaphical interpertation

d)	Compute and plot the seasonally adjusted data. 
```{r plotseasonal, echo=FALSE}
visitor.seasonal = stl(visitors, s.window="periodic")$time.series[,"seasonal"]
plot(visitor.seasonal)
```

#e)	Change one observation to be an outlier (e.g., add 500 to one observation), and recompute the seasonally adjusted data. What is the effect of the outlier? 
```{r outlier1, echo=FALSE}
visitor.seasonal[5] = 500
plot(visitor.seasonal)
```
##The effect is that there is a large value that clearly is an outlier and skews the graph

f)	Does it make any difference if the outlier is near the end rather than in the middle of the time series? 
```{r outlier2, echo=FALSE}
visitor.seasonal[230] = 500
plot(visitor.seasonal)
```
##No difference other than the outlier is on the other end of the chart

g)	Use STL to decompose the series. 
```{r STL, echo=FALSE}
visitors.stl = stl(visitors, s.window = "periodic")
plot(visitors.stl)
```
